### 2020-07-29 MSc_Dissertation_B156476.tex
Substantial additions * More explanations / observations in results * Added Jaccard heatmap of candidate gene set similarities * Good start at Discussion, analysing each component.

% MSc Dissertation on pattern expression in Ovarian Cancer

\documentclass[draft, tikz, 12pt,a4paper,oneside,fleqn]{article}
\usepackage[margin=1.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{setspace}
\usepackage{enumitem}
\setlist[description]{leftmargin=\parindent,labelindent=\parindent}

\usepackage{caption}
\captionsetup[figure]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\captionsetup[table]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{afterpage}

\usepackage[linewidth=1pt]{mdframed}

\graphicspath{{/home/ipoole/Documents/gitrepos/MScDissertation/images/}}  % trailin '/' is required!
\newcommand{\plotspath}[1]{{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/#1}}

\definecolor{darkblue}{rgb}{0.0, 0.0, 0.55}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=darkblue,
    filecolor=blue,      
    urlcolor=darkblue,
}
\usepackage{subfig}
\usepackage{booktabs}  % Needed for pandas to_latex() output
\usetikzlibrary{positioning}

\newcommand{\etal}{{\em et al\/}}
\newcommand{\comment}[1]{\marginpar{{\tiny\singlespacing #1 \par}}}

\linespread{1.5}
% This sets Arial font (bit ugly!)
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}
\setlength{\parindent}{2em}
\setlength{\parskip}{1em}
\renewcommand{\baselinestretch}{1.5}

%\usepackage{titlesec}
%\titleformat{\subsubsection}
%   {\itshape\normalsize}{\thesubsubsection}{1em}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% SUBMISSION PAGE %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\thispagestyle{empty}

\title{Phylogenetic study of Asgard Archaea}
\author{}

\begin{document}

%%%%%% Assesment Title Page
\begin{center}
\includegraphics[scale=0.3]{images/UoE_SBO_logo.png}
\end{center}

\vspace{0.3in}

\begin{mdframed}
\begin{center}
\huge
\vspace{0.3in}
\bf
Disentangling patterns of gene expression in high grade serous ovarian cancer
\vspace{0.2in}
\end{center}
\vspace{0.2in}
\end{mdframed}

\vspace{0.3in}

\begin{mdframed}
\begin{center}
\large
\vspace{0.2in}
Student Exam Number: \bf{B156476}
\vspace{0.2in}
\end{center}
\end{mdframed}

\vspace{0.3in}

\begin{mdframed}
\begin{center}
\large
\vspace{0.2in}
In partial fulfilment of the requirement for the Degree of
Master of Science in Systems and Synthetic Biology at the
University of Edinburgh,
2019 / 2020
\vspace{0.2in}
\end{center}
\end{mdframed}

\vspace{0.3in}

\begin{mdframed}
\begin{center}
\large
\vspace{0.2in}

Disssertation Supervisor:  Dr. Ailith Ewing
\vspace{0.2in}
\end{center}
\end{mdframed}

\newpage
\setcounter{page}{1}
%\pagestyle{headings}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Start of report for real %%%%%%%%%%%%%%%%%%%

\section{Introduction}
\emph{Say something about ovarian cancer!   An image perhaps.}

\subsection{Matrix factorization methods}
\label{sec-matrix-factorization-intro}

A good introduction to field is ref \cite{Stein-OBrien2018}, in which matrix factorization (MF) methods are discussed as a means of extraction the low dimensional structure from a high dimensional ``-omics" datasets, with gene expression analysis being the prime example.

There are three matrix factorization methods of interest to this project:
\begin{enumerate}
\item Independent component analysis (ICA)
\item Non-negative matrix factorization (NMF)
\item Principal component analysis (PCA)
\end{enumerate}

Conventionally, transcriptomics expression arrays are oriented with genes (or transcripts) as rows, and samples (e.g. patients) in columns.   
A typical expression array might have in the order of tens of thousands of rows and hundreds of columns.  
MF methods reduces this large $M \times N$ (genes $\times$ samples) matrix into two two smaller matrices.
In the general terminology of Stein-O'Brien \etal\ these are the $M \times K$ \emph{pattern matrix} and the $K \times N$ \emph{amplitude matrix}.  $K$ is typically small of the order 10, and refers to the number of extracted \emph{factors} (or components). 
Since we are focussing on transriptomics, rather than pattern and amplitude we will use the terms \emph{metagenes} and \emph{metasamples} respectively, which are in common useage.  The symbol conventions for these matrices varies depending whether NMF, ICA or PCA is being discussed.   NMF generally uses $W$ and $H$ respectively while ICA generally uses $S$ and $A$.  PCA is tends to be differently formulated, but $W$ and $T$ sometimes used.   The original matrix to be factorized is variously named $X$ or $V$.

In this work we wish to treat and discuss the three factorization methods in a unified way.  Thus, for the remainder of this dissertation we will adopt the following conventions:
\begin{description}
\item[Expression matrix:] $X$ $(M \times N)$
\item[Metagene matrix:] $W$ $(M \times K)$
\item[Metasample matrix:] $H$ $(K \times N)$
\end{description}
where $N$ is the number of patients (or samples), $M$ is the number of genes and $K$ is the factorization rank, i.e. the number of components or factors.  Thus the factorization is written
\begin{equation}
X \approx W H
\end{equation}
Adding subscripts to indicate the row,column orientation of the matrices:
\begin{equation}
X_{M,N} \approx W_{M,K} H_{K,N}
\end{equation}
as illustrated in figure \ref{fig-matrix-factorization} (a).

\begin{figure}[ht]
\def\M{2.2cm}
\def\N{1.3cm}
\def\K{0.3cm}
\def\fudge{-0.8mm}
\centering
\subfloat[Optimising independence of Meta\underline{genes}]{
\begin{tikzpicture}[node distance = 0.25cm]
	\node (X) [draw, fill=green, minimum height=\M, minimum width=\N] {\large X};
	\node [right=of X] (eq) {\large $\approx$};
	\node [right=of eq] (S) [draw, fill=yellow, minimum width=\K, minimum height=\M] {\large W};
	\node [right=of S] (times) {\LARGE$\times$};
	\node [right=of times] (A) [draw, fill=cyan, minimum width=\N, minimum height=\K] {\large H};

	\node [below=\fudge of X] {\tiny{$N$ Samples}};
	\node [left=\fudge of X] {\rotatebox{90}{\tiny{$M$ Genes}}};
	\node [below=of X, text width=2cm, align=center] {\small Expression matrix};

	\node [below=\fudge of S] {\tiny{$K$ Metagenes}};
	\node [left=\fudge of S] {\rotatebox{90}{\tiny{$M$ Genes}}};
	\node [below=of S, text width=2cm, align=center] {\small Metagenes};

	\node [below=\fudge of A] {\tiny{$N$ Samples}};
	\node [left=\fudge of A] {\tiny{$K$}};
	\node [below=of A, text width=2.5cm, align=center] {\small Metasamples};
\end{tikzpicture}}

\subfloat[Optimising independence of Meta\underline{samples}]{
\begin{tikzpicture}[node distance = 0.25cm]
	\node (X) [draw, fill=green, minimum height=\N, minimum width=\M] {\large X};
	\node [right=of X] (eq) {\large $\approx$};
	\node [right=of eq] (S) [draw, fill=cyan, minimum width=\K, minimum height=\N]{\large W};
	\node [right=of S] (times) {\LARGE$\times$};
	\node [right=of times] (A) [draw, fill=yellow, minimum width=\M, minimum height=\K]{\large H};

	\node [below=\fudge of X] {\tiny{$M$ Genes}};
	\node [left=\fudge of X] {\rotatebox{90}{\tiny{$N$ Samples}}};
	\node [below=of X, text width=2cm, align=center] {\small Expression matrix};

	\node [below=\fudge of S] {\tiny{$K$ Metasamples}};
	\node [left=\fudge of S] {\rotatebox{90}{\tiny{$N$ Samples}}};
	\node [below=of S, text width=2.5cm, align=center] {\small Metasamples};

	\node [below=\fudge of A] {\tiny{$M$ Genes}};
	\node [left=\fudge of A] {\tiny{$K$}};
	\node [below=of A, text width=2cm, align=center] {\small Metagenes};
\end{tikzpicture}}
\caption{Two ways of configuring matrix factorization in the context of gene expression analysis.
In configuration (a) NMF and ICA will optimise the \emph{metagenes}, while in (b) the \emph{metasamples} are optimised. }
\label{fig-matrix-factorization}
\end{figure}

However, in the case of NMF and ICA, the $W$ and $H$ matrices cannot be trivially exchanged and transposed.  This is because the optimisations (sparsity and independence respectively) which define these algorithms are focussed on the $W$ matrix.  It is perfectly possible to apply these algorithms to gene expression analysis with exchanged and transposed meanings of $W$ and $H$, and this is illustrated in figure \ref{fig-matrix-factorization} (b).  In this case it is the properties of the \emph{metasamples} which will be optimised.  Thus, the two forms are different in substance, not simply in notational convention.

There seems to be substantial confusion an lack of clarity in the way that matrix factorization, particularly ICA, is applied in transcriptomics research.  
``Surprisingly, both ways of applying ICA to omics data are wide-spread, and sometimes it makes an effort to figure out in which way ICA was applied." \cite{Sompairac2019a}, and ``Different protocols to apply ICA to transcriptomic data exist and currently no single standard approach has been defined. The main difference in the existing approaches consists in what is considered as source signal matrix in the decomposition" \cite{Cantini2019}.  
According to Cantini \etal\, references \cite{Au-Yeung2014}, \cite{Kairov2017}, \cite{Kong2008} and \cite{Lee2003} optimise metagenes, while references \cite{Meng2016} and \cite{Barillot2013} optimises metasamples.

\subsubsection{Determining the optimum number of factors}

The number of factors, $K$, to extract is a key decision in any matrix factorization approach.  In this regard, PCA differs from ICA and NMF.  In PCA it is reasonable to extract all factors, setting $K=N$, the factors being ranked by the associated eigenvalue which also ranks the proportion of variance explained.   That approach is not reasonable for ICA and NMF, since differing sets of factors will be obtained for different $K$, and for a given choice of $K$ all factors are equally important -- they do not rank \cite{Stein-OBrien2018}.

The problem of determining the optimal $K$ for a given expression matrix is addressed  by Kairov \etal\ \cite{Kairov2017}, based on optimizing the \emph{stability} of the components over multiple runs.

\subsection{Ovarian cancer sub-type prediction}

A recent pre-publication develops a pipeline for sub-type prediction which the potential to be used clinically \cite{Talhouk2020}.
The NanoString gene expression platform is used. Approach is to focus on small set of 513 genes (vs 20,000 protein coding genes) known have relevance to subtyping, as identified from literature.

Two approaches were followed by different teams.  ``All Array": used expression array data from 1650 patients across 14 studies evaluating 9 supervised learning algorithms by bootstrap, selecting an AdaBoost -like method. ``TCGA" used 434 patients from TCGA evaluating 5 algorithms by cross-validation selecting a random forest.

"All Array" had the advantage of more data but needed attention to batch effects. A final classifier took a consensus of the two methods based on a minimal gene set (order 40 genes), validated by a leave-one-out (patient level) approach.

Survival analysis was carried out stratified by predicted subtype.

\begin{figure}
\begin{center}
\includegraphics[scale=0.35]{talhouk_figure_1_part.png}
\end{center}
\caption{HGSOC subtype prediction by expression analysis of a focussed gene set and concensus across two independent pipelines.  From \cite{Talhouk2020}.}
\label{fig-talhouk_figure_1_part.png}
\end{figure}

\subsection{Gene expression in HGSOC}

This section focusses on biological and clinical conclusions relating to gene expression in HGSOC and mentions analysis techniques only in passing.

Obesity is known to negatively impact prognosis in ovarian cancer \cite{Cuello2018}, \cite{Au-Yeung2014}.  Cuello \etal\ \cite{Cuello2018} used NMF based clustering on mRNA microarrays and reverse-phase protein arrays to demonstrate an association between cancer driver genes and obesity related genes.

Wang \etal\ \cite{Wang2017c} also use NMF clustering to identify five subtypes of HGSOC informative of outcomes: 1:mesenchymal, 2:immunoreactive, 3:proliferative, 4:differentiated and 5:anti-mesenchymal.  Subtypes 2, 5 are found to be associated with longer survival.

Mairinger \etal\ screened 770 immune related genes to identify 11 differentially expressed genes associated with response to platinum treatment.  They find that expression of HS11B1, DNBT1, CKLF, NUP107. CCL18, LY96, ATG7, SLAMF7, CXCL9 is associated with long survival, while IKBKG and SDHA with short survival.

BRCA1 inactivation is known to cause chromosomal instability in many cancers.  Pradhan \etal\ \cite{Pradhan2010} investigated the role of BRCA1 in HGSOC by copy number and expression analysis, finding surprisingly that inactivation has no relationship with gross genomic alteration.  They leave open the question whether DNA repair by PARP plays a role.   The relationship between BRCA1/2 and DNA repair -- specifically homologous repair deficiency (HRD) -- is picked up by Ewing \etal \cite{Ewing2020}.  They study the complex interplay of single nucleotide variants and large structural variants as they impact BRCA1/2 disruption and thence HRD.  This work suggests that when BRCA1/2 loss is detected in HGSOC patients, there may be a clinical role for PARP inhibitors, since this would prevent error-prone non-homologous repair by PARP, and so selectively kill BRCA1/2 disrupted cells.

Note that several of the studies reviewed here use data from the Australian Ovarian Cancer Study (AOCS): \cite{Winterhoff2016,Patch2015,Ewing2020,Zhang2018,Cuello2018,Au-Yeung2014}.

Stratification of HGSOC by immunohistochemistry is addresses Kamble \etal. \cite{Kamble2019}.  They identify a panel of six biomarkers -- TCF21, E-cadherin, PARP1, Slug and AnnexinA2.   These markers allow stratification into categories including Mesenchymal-to-epithelial transition, Homologous recombination repair and Epithelial-to-mesenchymal transition.  This, they argue, will allow selection of class-specific inhibitor drugs such as Olaparib and Rucaparib.

The platinum based drug Cisplatin is a key treatment in ovarian cancer, yet tumours often develop resistance, possibly related to the host's immune response.  Understanding and predicting such resistance is therefore of clinical importance and is addressed by Maringer \etal\ \cite{Mairinger2019} through expression analysis using the NanoString platform with a panel of 770 immune related genes.   They find the following to be significantly related to platinum resistance: KLRC1, TCF7, CD274, HSD11B1, COLEC12, PDGFC, FCF1, BMI1, TNFRSF9, ATG10, EWSR1.

\subsection{Research questions}
The following research goals were set out early in project. 
\begin{enumerate}
\item Which are the influential genes in HGSOC according to gene expression data?   Do these confirm published results?
\item Which of PCA, ICA and NMF is best suited to this analysis?
\item For identified genes, what is the underlying biology?
\item Can we discriminate subtypes of HGSOC from expression data?
\item Is there value in incorporating somatic and germline genomic data?
\end{enumerate}

%%%%%%%%%%%%%%% METHODOLOGY %%%%%%%%%%%%%%%%%

\section{Methodology}

\subsection{Outline}

Our overall approach is to use \emph{unsupervised} machine learning methods to represent gene expression data in a small number of features, then to study whether these features correlate with clinical and biological variables.  This is appropriate in our case, since of the two datasets available (described below), one (TCGA) has relatively large n but little available metadata, while the other (AOCS) is much smaller but has useful metadata.   This motivates an approach of unsupervised learning on the TCGA dataset.

\subsubsection{Datasets}
Two gene expression datasets were assailable for this work (n refers to number of patients, g refers to number of genes).
\begin{enumerate}
\item TCGA, n=274, g=19,601, with metadata on survival
\item AOCS, n=80, g=19,730, with metadata on survival, cellularity and additional genomic features
\end{enumerate}

\subsubsection{Method outline}
An overview of the methodology adopted in this work can be seen in figure \ref{fig-method-overview}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.7, draft=false]{methodlology_overview.pdf}
\end{center}
\caption{Overview of methodology as information flow}
\label{fig-method-overview}
\end{figure}
In outline our method is follows:
\begin{enumerate}
\item
Identification of a consistent gene list between the TCGA and AOCS datasets.
\item
Unsupervised metagene extraction by matrix factorization methods : NMF, ICA and PCA.
\item
Metagene selection for robustness by k-means clustering of bootstrap resampled factorizations evaluated by silhouette score, based on the TCGA dataset.
\item
Investigation of biological significance of determined factors by gene enrichment analysis against the Gene Ontology (GO).
\item
Computation by least squares optimisation of metasamples associated with with each TCGA derived metagene against TCGA and AOCS datasets.
\item
Alignment of available metadata with computed metasamples, for TCGA and AOCS.
\item 
Survival analysis on TCGA and AOCS to investigate relationship between metasamples and patient survival.
\item
Clustered heatmaps to visualise relationships between metasamples, patients and metadata.
\end{enumerate}

\subsubsection{Tools}
\begin{itemize}
\item Python 3.6
\item PyCharm IDE
\item Numpy for high performance matrix manipulation
\item Pandas for data frame handing
\item Matplotlib for general plotting
\item Seaborn for clustered heatmaps
\item Scikit-learn for matrix factorization and k-means clustering
\item GOATOOLS package for gene enrichment analysis against GO
\item Lifelines package for survival analysis.
\end{itemize}

A full list of software and libraries used, with versions, is given in appendix \ref{sec-software-versions}

\subsection{Data normalization and quality control}

Expression data for the AOCS and TCGA datasets was received for this project in a spreadsheet format, having been derived from the RNA-Seq data as described in \cite{Ewing2020}, including  normalization by variance stabilizing transformation.

\subsection{Gene set intersection}
\label{sec:gene-set-intersection}

In order to allow factorizations found in the TCGA dataset to be applied to the AOCS dataset it is necessary (or at least convenient) to synchronize the set of genes over which the expression matrices are defined.   As provided to this project, the TCGA and AOCS datasets cover 19,610 and 19,730 protein coding genes respectively, 19,566 or which common to both (according to ENSG encodings).  Thus, both datasets were pruned to the 19,566 intersection set and ordered consistently.

\subsection{Matrix factorization computation}
One of the aims of this work is to compare the efficacy of three methods of dimensionality reduction -- NMF, ICA and PCA -- as explained in the introduction.   The three methods have different properties and notational conventions, but to simplify the discussion here we adopt the notation $X \approx W H$ for all three methods, where $X$ is the (genes, patients) expression matrix  $W$ is the (genes, factors) metagenes matrix and $H$ is the (factors, patients) metasamples matrix.

Some of the key algorithm hyper parameters of each method were investigated and tuned with respect to reconstruction accuracy, specifically the root-mean square (RMS) difference between $X$ and $W H$. The following parameters were explored in each case:
\begin{description}
\item[NMF:] Parameters {\tt max\_iter} (algorithm iterations) and {\tt tol} (tolerance of convergence) were optimised for good accuracy and acceptable execution time.  Other parameters of interest are {\tt alpha} (multipluer for regulation term) and {\tt l1\_ratio} (multiplier for L1 regularization, when {\tt alpha} $> 0$).  L1 regularization  favours elements being precisely zero, whereas L2 regularization will encourage them to be small.  These parameters have not been explored in this project, and the default {\tt alpha = 0} (no regularization) was used.
\item[ICA:]  Parameters {\tt max\_iter} and {\tt tol} were investigated as for NMF above.  Additionally, options for the entropy functional which forms the basis of the optimiziation were investigated.
\item[PCA:] This is in principle a deterministic algorithm based on eigenvector decomposition.  However, {\tt sklearn.decomposition.PCA} uses a more efficient 'randomized' algorithm when the given matrix is larger than 500 in both dimensions.  Thus in our use case PCA is seen to have stochastic behaviour.

\end{description}

\subsection{Metagene selection by cluster coherence}
\label{sec-metagene-selection}
As outlined in the discussion, deciding on the number of components (metagenes) to extract -- that is the factorization rank -- is key.  Taking more components results in more accurate representation of the observed expression matrix and provides more avenues to explore the underlying biology.  However, it is important that the metagenes are \emph{stable}, that is that they have reliable meaning when transfered to other datasets.   There are two sources of variation or instability to consider.

Firstly, NMF and ICA are inherently stochastic algorithms, sensitive to their starting state, so repeated runs give different results.

A second and more fundamental source of variation of relevance to all three factorization methods is \emph{sampling error}.   Our factorizations are based on a small (N=80 or N=374) sample of patients drawn from the population of HGSOC patients; our particular datasets are just two examples of many different `draws' which could have been made from that population.

\emph{Bootstrap sampling} (also know as \emph{Monte Carlo} simulation) is a common method of empirically propagating the consequence of sampling error when the distribution or processing operations are difficult to model mathematically.   This is implemented by performing factorizations multiple times, at each iterations choosing N samples from the N available \emph{with replacement}.   For this work 50 repeats were performed, being a compromise between achieving an adequate simulation without overly burdensome computation time.  The overall process is summarised as pseudocode in figure \ref{fig-clustering-psuedocode}.

Each of the three factorizer methods was evaluated for rank $k$ between 2 and 10.  In each case, 50 iterations of factorization are performed on bootstrap samples, generating $50 k$ instances of 19,566 dimensioned metagenes.  Two dimensional t-SNE plots \ref{tsne-reference} were generated for visualization.   If sampling and algorithm initialisation error are modest then we expect to see $k$ clusters of points.   To avoid lengthy computation in the t-SNE clustering, the metagenes were first reduced to $r=20$ dimensions by PCA; brief experiments showed that this reduction had negligible effect on the t-SNE visualisation for $r \geq 10$.

In order to obtain median estimates of the $k$ metagenes from the $50 k$ which were generated, k-means clustering was performed in the PCA reduced (20 dimension) space, delivering $k$ sets of points.  These were referenced back to the original 19,566 dimensioned metagenes and the per-dimension median calculated.  These $k$ median metagenes were saved to a file named for the specific factorizer and rank $k$.

The k-means clustering further allowed for a quantitative assessment of cluster coherence for the particular choice of factorizer and rank, via the \emph{silhouette score}.  In brief, this is a measure of how close (by Euclidean distance) each point in a cluster is to other points in the cluster versus points \emph{not} in the same cluster.  See 
\href{https://en.wikipedia.org/wiki/Silhouette_(clustering)}{Wikipedia: Silouette (clustering)}.
The score is in the range -1 to +1, with 0 implying random scatter and 1 implying all points of a cluster perfectly overlay.

The t-SNE plots and silhouette scores were assessed visually (see figures \ref{fig-first-cluster-result} to \ref{fig-last-custer-result} to decide, for each of the three factorizers, the highest rank with good cluster coherence, those median metagenes to take forward to gene enrichment and survival analysis.   
Note that the t-SNE plots are used only for visualisation; they are not involved in the computation of median metagenes or silhouette score.

Initial investigation on the N=80 AOCS dataset showed very poor cluster coherence, even for k=2.   Thus it was decided to compute and select metagenes only from the N=374 TCGA dataset.   
The selection rationale is set out in the results section, but is convenient to state here that the following ranks were selected:
$K_{\mbox{\tiny NMF}} = 3, K_{\mbox{\tiny ICA}} = 5, K_{\mbox{\tiny PCA}} =3$.  Thus, there were $3 + 5 + 3 = 11$ metagenes taken forward for follow on analysis.

\begin{figure}
\begin{center}
\begin{Verbatim}[baselinestretch=1, frame=single, rulecolor=\color{blue}, label=Metagene Stability Assessment, fontfamily=courier, fontsize=\small]

 for factorizer in NMF, ICA, PCA:
    for k in 2..10:
       Ws = []
       for j in 1..50:
          X = bootstrap sample N from N
          W, H = factorizer(X, rank=k, seed=j)
          append W to Ws

          # Ws is a list of 50*k metagenes, each of length 19,566
          # Pragmatically reduce metagenes to 20 dimensions by PCA

       Ws_reduced = PCA(Ws, rank=20)
       plot t-SNE(Ws_reduced)
       clustering = k-means clustering(Ws_reduced, n_clusters=k)
       score[factorizer, k] = silhouette score(clustering)	  
       median_metagenes = calculate medians for k clusters of metagenes
       save median_metagenes to file by factorizer and k
       for each factorizer 
          plot score vs k

\end{Verbatim}
\end{center}
\caption{Pseudocode for generating median metagenes and assessing there stability to random algorithm initialization and sampling error.}
\label{fig-clustering-psuedocode}
\end{figure}

\subsection{Gene enrichment analysis}

The metagenes extracted by the above described factorization process provide valuable information into which genes varying in expression in the study samples; thus in our case, which genes are influential in HGSOC.

In order to gain insights into what biological processes are involved, gene enrichment analysis against the Gene Ontology (GO) was carried out for each of the 11 metagenes individually, as follows.

The essence of gene enrichment analysis is to compare a candidate set of genes with many functional gene sets, asking the question "are there significantly more (or less, i.e. depletion) intersecting genes than would be expected for the same number of genes being drawn at random (without replacement) from the total population of genes in the study.

For each metagene, the candidate gene set was determined as those genes having weights outwith three standard deviations of the mean.   This set was analysed using the Python GOATOOLS package \cite{Klopfenstein2018}.

The gene ontology was downloaded from \url{http://purl.obolibrary.org/obo/go/go-basic.obo}.  ({\tt Purl.org} is a resource for managing permanent URLs; {\tt obo} refers to the Open Biological and Biomedical Ontology (OBO)).  
Annotations linking human genes to GO concepts was downloaded from the GO website, specifically \url{http://geneontology.org/gene-associations/goa_human.gaf.gaf}.   
The gene population was defined as the common 19,566 protein coding genes (see section \ref{sec:gene-set-intersection}) against which the metagenes were computed.

Uncorrected p-value threshold was set to 0.01.
Multiple hypothesis significance testing used the false discovery rate (FDR) method of Bengamini and Hochberg, the FDR threshold being set to 0.01.  This is more stringent than the 0.05 value which is typically used, and was chosen since multiple metagenes are being analysed, implying multiple hypothesis testing over and above that which is accounted for by FDR filtering within a single gene enrichment analysis run.

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.13]{GO_0042773_lineage.png}
\caption{Example of a small section of the Gene Ontology (GO), focussed GO:0042773 (in purple), showing parents and children of the term.   Generated by GOATOOLS.}
\label{fig:eg_go_graphic}
\end{center}
\end{figure}

The result of this analysis, per metagene, is a list of enriched (or depleted) GO terms, each with an associated list of involved genes and a FDR significance level.   This list can be inspected directly for insights, but that misses the point of the GO, which is to organise the terms hierarchically by 'is a' relationships.   Thus, the enriched terms were rendered graphically to show them in the context of their parent terms.

\subsection{Transfer of learned metagenes to novel a dataset}
\label{sec-transfer-to-novel}

\newcommand{\trainset}{\mathcal{T}}
\newcommand{\validset}{\mathcal{V}}

It is fundamental to our approach that metagenes determined on the basis of one cohort of data can be used to generate metasamples for a different cohort.  In the exposition below we refer to these cohorts as the \emph{training} and \emph{validation} cohorts respectively.

Rank $k$ matrix factorization on the training dataset $\trainset$ (e.g. TCGA) results in:
\begin{equation}
 X_\trainset \approx W_\trainset H_\trainset
\end{equation}
where $X_\trainset$ is the expression matrix of shape $(g_\trainset, n_\trainset)$, with $g_\trainset$ the number genes and $n_\trainset$ the number of patients. $W_\trainset$ is the metagene matrix of shape $(g_\trainset, k)$ and $H_t$ is the metasamples matrix of shape $(k, n_\trainset)$.

We wish to apply the factorization learned on $\trainset$ to a novel dataset $\validset$ (specifically the AOCS dataset) of shape $(g_\validset, n_\validset)$. 
Importantly, $n_\validset = 1$ reflects the application of these methods to a single patient in a clinical setting.  To apply the learned factorization we need to find $H_\validset$ as in the factorization
\begin{equation}
	X_\validset \approx W_\validset H_\validset
\end{equation}
In both the experimental and clinical situation we are given $X_\validset$ but not $W_\validset$ and require to find $H_\validset$. 
We \emph{cannot} simply perform the matrix factorization on dataset $\validset$ since $n_\validset$ may be small, or even a single patient.   
However, if patients in datasets $\trainset$ and $\validset$ are drawn from the same population (ovarian cancer patients), then $X_\trainset$ and $X_\validset$ can be expected to have similar distributions w.r.t to their columns, and thus $W_\trainset$ and $W_\validset$ can be expected to be equivalent within sampling error.  
This only makes sense if the two expression matrices $X_\trainset$ and $X_\validset$ are defined over the \emph{same list of genes}, so that $g_\trainset = g_\validset = g$, in which case $W_\trainset$ and $W_\validset$ have the same shape of $(g, k)$.

We thus need to solve for $H_\validset$ in
\begin{equation}
	X_\validset  \approx  W_\trainset H_\validset \label{eq_Xv_WtHv}
\end{equation}

This can be solved by the method of least squares.  In the case that the original factorization was by NMF, we use non-negative least square regression (NNLS):
\begin{equation}
	H_\validset = \mbox{argmin}_{H\geq 0}{\lvert W_\trainset H - X_\validset \rvert}_2
	\label{eq_Hv_WdtXv}
\end{equation}
where ${\lvert \mathbf{\cdot} \rvert}_2$ indicates Euclidean distance or L2-norm.
The Python library function {\tt scipy.optimize.nnls} is used.

For ICA and PCA we use ordinary least square regression, formulated above by dropping the $H\geq 0$ constraint, using {\tt scipy.linalg.lstsq}.

An internal check on the validity of the statistical assumptions can be made my substituting  $H_\validset$ from (\ref{eq_Hv_WdtXv}) into (\ref{eq_Xv_WtHv}) and assessing the accuracy of the approximation by RMS difference.

The end result of the analysis is the $H_\validset$ matrix of shape $(k, n_\validset)$, thus delivering for dataset in our $\validset$ a feature vector of length $k$ -- a small number typically less than a dozen.

Since in this work we wish to evaluate the efficacy of the three factorization methods -- NMF, ICA and PCA -- we in fact take forward three different $H$ matrices, each having ranks as for the associated metagenes.

\subsection{Reconciling computed metasamples and metadata}
\label{sec-reconciling-metasamples-metadata}

To carry out patient level analysis -- survival analysis, heatmaps and boxplot -- as described in the following sections, it is necessary to compile tables (Panda DataFrames) which bring together per-patient metasamples relating to each of the selected metagenes, with associated metadata -- such as cellularity and survival information.   The metagenes always derive from the TCGA dataset (because of its larger size as explained in section \ref{sec-metagene-selection}), but we wish to apply these metagenes to the expression data and associated metadata of either the TCGA or AOCS datasets.   
The mathematics of transferring metagenes across datasets has been described above.  Care is required to ensure that expression matrices aligned correctly with metagenes with respect to their genes (rows), and with the metadata with respect to patient identifiers in the columns.
Note that the same transferring approach is taken when metasamples are required for TCGA, even though we could in that case obtain the $H$ matrix directly from the factorization.

\subsection{Survival analysis}

Survival analysis was performed to investigate whether the selected metagene factors correlate with patient survival.  Overall  survival (OS) data is available for both the TCGA and AOCS datsets.  Additionally, progression free survival (PFS) data is available for AOCS.   As discussed in section \ref{sec-clustering-method}, metagenes were obtained by factorization on the TCGA dataset only.  Application to TCGA (for OS) thus represents an in-sample test, while application to AOCS (for OS and PFS) is a more exacting out-of-sample test. There are thus three analyses to consider: 1) TCGA$\rightarrow$TCGA (OS), 2) TCGA$\rightarrow$OACS (OS) and 3) TCGA$\rightarrow$OACS (PFS).

Analysis was performed using the Python {\tt Lifelines} package \cite{Davidson-Pilon2020}.  %Metasamples ($H$ matrices) were calculated from the metagenes and expression matrix as explained in section \ref{sec-transfer-to-novel} below.  
Metasample values were binariased to 0, 1 by thresholding at the median value.   Kaplan-Meier plots were made in respect of derived metasample for each of the three analysis -- making $11 \times 3$ plots each with two survival curves (with 95\% confidence intervals).  A hazard ratio (HR)
\footnote{Do I need to explain the definition and meaning of HR?}
was calculated for each plot by fitting Cox's proportional hazards model with p-value relating to the hypothesis that HR is significantly different to 1.0.

\subsection{Metasample heatmap analysis}

To further investigate the relationship between metasamples and to available patient metadata, clustered heatmaps were generated using the \href{https://seaborn.pydata.org/generated/seaborn.clustermap.html}{Seaborn {\tt clustermap()}} function from the reconciled metasamples and metadata (section \ref{sec-reconciling-metasamples-metadata}).  Heatmaps were generated for TCGA and AOCS datasets (\emph{Should say something about stratifying the metadata, when I figure it out}).  Heatmap clustering by correlation distance was applied.

\subsection{Codebase}

The described method was implemented in Python.
%except for the KEGG gene enrichment analysis which was implemented in R.
All code is available in a github repository: \url{https://github.com/ipoole/HgsocTromics}.  Good software engineering practice has been followed, with an object-oriented design, frequent commits and unit testing.  Unit tests are based on tiny expression matrices, just 100 genes by 10 patients. The whole test suite of over 60 tests executes in around 30 seconds.  These tests are particularly valuable when re-factoring code by providing a rapid means of detecting errors.  All plots are generated in vector graphics pdf format to ensure smooth scaling to any resolution.

%%%%%%%%%%% RESULTS %%%%%%%%%%%%%
% \afterpage{\clearpage}

\section{Results}

Results of applying the above described methodology are explained and presented here.   Deeper interpretation and discussion is deferred to the following section.  Throughout, consistent colours of \textcolor{blue}{blue}, \textcolor{orange}{orange} and \textcolor{green}{green} are used for results relating to NMF, ICA and PCA respectively.  Specific metagenes are referred to by, for example, ``NMF-2-of-3" -- meaning the 2nd component of the rank $k=3$ NMF factorization.

% \subsection{Investigation of optimal parameters}

\subsection{Cluster coherence analysis results}

From figure \ref{fig-AOCS-ica-fixed-vs-bootstrap} it can be seen that demonstrates that with sampling error excluded, clusters appear reasonably coherent, but when sampling error is modelled by bootstrap sampling then we see the factorizations become much less stable.  This demonstrates that \emph{sampling} error is far greater then errors due to algorithm initialisation when the dataset is small (N=80), making metagene extraction highly unstable.   Figure  \ref{fig-TCGA-ica-fixed-vs-bootstrap} makes the same comparison for the lager N=374 TCGA dataset, in which it can be seen that the impact of sampling error is not so severe.  For brevity only results for ICA are shown, but a very similar effect is observable for PCA and NMF.

For this reason, and in line with our overall strategy, it is appropriate to perform all metagene extraction on the larger TCGA dataset, in the expectation that the obtained metagenes will be more robust an likely to generalise better to other datasets.

\def\S{0.4}
\begin{figure}
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_ICA_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_ICA_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from ICA facorizations on the N=80 AOCS dataset, comparing \emph{fixed} sampling (top row) with \emph{bootstrap} sampling (bottom row) for a selection of factorization ranks.  }
\label{fig-AOCS-ica-fixed-vs-bootstrap}
\label{fig-first-cluster-result}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_fixed.pdf}} 
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_bootstrap.pdf}}
\caption{Clustering of metagenes from ICA facorizations on the N=374 TCGA dataset, again comparing fixed and bootstrap sampling.}
\label{fig-TCGA-ica-fixed-vs-bootstrap}
\end{center}
\end{figure}

The critical decision of the choosing the factorization rank for each method -- $K_{\mbox{\tiny NMF}}$, $K_{\mbox{\tiny ICA}}$ and $K_{\mbox{\tiny PCA}}$ -- is based on figure \ref{fig-TCGA-all-bootstrap}.   On this bases, considering both the t-SNE plots and silhouette scores, the following choices were made:
$K_{\mbox{\tiny NMF}}  =  3,
K_{\mbox{\tiny ICA}}  =  5,
K_{\mbox{\tiny PCA}}  =  3$.

The rationale for these choices is that, firstly $k>2$ is desirable to have sufficient information to work with.  NMF cluster coherence seems to deteriorate after $k=3$.   Looking at the graph of silhouette scores (figure \ref{fig-TCGA-all-bootstrap}, bottom), ICA appears to have a sweet spot at $k=5$.  For PCA, $k = 3$ or $4$ both seem reasonable; $k=3$ was selected.

There is a curious artefact visible in the NMF factorization clusterings of figure \ref{fig-TCGA-all-bootstrap}, top row.  For $k=2, 3 and 5$ several of the clusters show a bi-modal character.  This is not observed in the fixed sampling case (not shown).  The artefact is difficult to explain.

\begin{figure}
\begin{center}
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_NMF_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_PCA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=0.55]{\plotspath{/TCGA_OV_VST/FactorClustering/silhouette_plots_TCGA_2_3_4_5_6_7_8_9_10_bootstrap.pdf}} 
\caption{Metagene clustering for all three methods applied to the N=374 TCGA dataset with bootstrap sampling, over a range of factorization ranks.  The silhouette scores are also plotted (bottom).  It is on the basis of this figure that the factorization ranks for each method were selected.}
\label{fig-TCGA-all-bootstrap}
\label{fig-last-cluster-result}
\end{center}
\end{figure}

\afterpage{\clearpage}

\subsection{Gene Enrichment (against GO) Results}

Gene enrichment analysis against the GO results in, each of the 11 metagenes, a table of GO terms with a list of the candidate genes which have an annotated association to that term.   
To take full advantage of the hierarchical nature of the GO, results are presented as lineage maps in figures \ref{fig-NMF-go-lineage} to \ref{fig-PCA-1-3-go-lineage}.  
The contributing candidate genes for all significantly enriched terms are shown beneath each figure.  
(Ideally, the specific genes would be listed alongside each term, but this is difficult to achieve graphically).   
High-level terms at depth less than 3 were removed, since these are generic (e.g. "regulation of biological process") and so uninteresting.   
One component -- ICA-4-of-5 -- was problematic in that 49 GO terms (with depth $\geq 3$) were identified as significant.  
For this component it was necessary to limit the graphic to the 12 terms having the largest number of associated candidate genes.   
The full table of terms for this component is available in appendix \ref{tab-ica-4-of-5-enrichment-table}.

The four components NMF-3-of-3, ICA-1-of-5, ICA-2-of-5 and PCA-2-of-3 yielded no significant enriched terms (for FDR $\leq$ 0.01).

\begin{figure}
\begin{center}
\includegraphics[scale=0.5, draft=false]{\plotspath{TCGA_OV_VST/GeneEnrichment/component_jaccard_heatmap.pdf}}
\end{center}
\caption{Heatmap of Jaccard similarities between the candidate genes identified by the 11 components.}
\label{fig-component-jaccard-similarities}
\end{figure}

To what extent do the 11 components overlap in the candidate genes they propose (independent of enrichment analysis)?  Figure \ref{fig-component-jaccard-similarities} shows this by a heatmap of Jaccard similarities.  Jaccard similarity (or index) is defined on a pair of sets, in this case sets of genes, as:
\[
   \mbox{Jaccard}(A, B) = \frac{A \cap B}{A \cup B}
\]
lying between 0 (no similarity) and 1 (identical).

From this heatmap we can see that within a factorization method the components have almost no similarity -- they represent largely independent projections, as one would expect.  Between methods, however, there is considerable overlap in some cases.  In particular PCA 2\_of\_3 vs ICA 1\_of\_5 (J=0.61), PCA 3\_of\_3 vs ICA 5\_of\_5 (J=0.49) and NMF 1\_of\_3 vs ICA 3\_of\_5 (J=0.39).

\begin{figure}[hb]
\begin{center}
\includegraphics[scale=0.14]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_1_of_NMF_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_1_of_NMF_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component NMF-1-of-3} \\
\vspace{1cm}
\includegraphics[scale=0.14]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_2_of_NMF_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_2_of_NMF_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component NMF 2-of-3} \\
\caption{Lineage maps of enriched Gene Ongology (GO) terms for components NMF-1-of-3 and NMF-2-of-3.  In these diagrams, enriched terms are coloured purple, while there ancestors in the ontology are yellow.  (NMF-3-of-3 produced no significant enrichment results)}
\label{fig-NMF-go-lineage}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.1]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_3_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_3_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component ICA-3-of-5} \\
\vspace{1cm}
\includegraphics[scale=0.07]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_5_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_5_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component ICA-5-of-5} \\
\caption{Lineage maps of enriched Gene Ongology (GO) terms for components ICA-3-of-5 and ICA-5-of-5}
\label{fig-ICA-3-5-go-lineage}
\end{center}
\end{figure}

\begin{sidewaysfigure}[ht]
\begin{center}
\includegraphics[width=\textwidth, draft=false]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_4_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_4_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
\caption{Lineage maps of enriched Gene Ongology (GO) terms for component ICA-4-of-5. For this component it was necessary to additionally filter the enriched terms -- see text.}
\label{fig-ICA-4-go-lineage}
\end{center}
\end{sidewaysfigure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_1_of_PCA_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_1_of_PCA_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component PCA-1-of-3} \\
\vspace{1cm}
\includegraphics[width=\textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_3_of_PCA_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_3_of_PCA_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component PCA-3-of-3} \\
\caption{Lineage maps of enriched Gene Ongology (GO) terms for components 1 and 3 from the rank=5 PCA factorization. (Component PCA-2-of-3 produced no significant enrichment results)  }
\label{fig-PCA-1-3-go-lineage}
\end{center}
\end{figure}

\afterpage{\clearpage}

\subsection{Survival analysis results}

Kaplan-Meier overall survival (OS) plots stratified by each metasample component are shown for the TCGA dataset in figure \ref{fig-kaplan-meier-os-TCGA-TCGA} and for the AOCS dataset in figure \ref{fig-kaplan-meier-os-TCGA-OACS}.  Plots for progression free survival (PFS) on OACS are shown in figure \ref{fig-kaplan-meier-pfs-TCGA-OACS}.   All plots show 95\% confidence intervals and hazard ratio (HR) with associated p-value.

All of these results are summarised in figure \ref{fig-summary-survival-plots}, which brings together the three sets of results -- TCGA (OS), OACS OS) and OACS (PFS) with respect to the 11 metasample components.  $\log_2{HR}$ is used in order that the sense of the survival impact can be readily appreciated.    If a metasample component has a robust correlation with survival, then we expect the p-value for all three sets of results to show significance \emph{and} for the effects to have the same sense -- be in the same direction.  None of the 11 components pass this test.

The component closest to correlation with survival would seem to be PCA 1-of-3, showing a reasonably consistent hazard ratio of around 1.3 ($2^{0.4}$) across the three experiments.  This has $p < 0.05$ in the larger TCGA dataset, but not in the smaller AOCS dataset.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_TCGA.pdf}} 
\caption{Kaplan-Meier plots for each metasample component, stratified at the median value, for TCGA $\rightarrow$ TCGA for overall survival (OS) case.  Hazard ratio and p-value is shown for each case.  The final plot (bottom right) is unstratified overall survival}
\label{fig-kaplan-meier-os-TCGA-TCGA}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for overall survival (OS).}
\label{fig-kaplan-meier-os-TCGA-OACS}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_pfs_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for progression free survival (PFS).  }
\label{fig-kaplan-meier-pfs-TCGA-OACS}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_NMF.pdf}} \\
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_ICA.pdf}} \\
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_PCA.pdf}}
\caption{Visual summary of survival analysis as applied to TCGA(OS), AOCS(OS) and AOCS(PFS).  Plots are divided by factorization method.  Bar heights show $\log_2(\mbox{HR})$ with p-value also shown.  For a component to have robust correlation with survival we expect all three of the analysis bars to pass a significance threshold {\em and} for the $\log{\mbox{HR}}$ to to be {\em in the same sense\/}.  None achieve this.}
\label{fig-summary-survival-plots}
\end{center}
\end{figure}

\afterpage{\clearpage}

\subsection{Metasample heatmap clustering results}

As with survival analysis, metasamples were derived from the metagenes factorized from the TCGA dataset, then applied to both TCGA and AOCS datasets.  The resulting heatmaps are shown in figures \ref{fig-heatmap-TCGA-TCGA} and \ref{fig-heatmap-TCGA-OACS} respectively.   Metadata was available for the AOCS dataset, and this is presented in colour coded columns on the left of figure \ref{fig-heatmap-TCGA-OACS}.   The meanings of each column are as follows:
\begin{description}
\item[WGD]: Whole genome doubling -- a marker of genome instability.
\item[Cellularity]: proportion of cells belonging to the tumour (as opposed to surrounding normal tissue).
\item[HR Detect]: A predictor of homologous repair deficiency based on established mutational features. \cite{Ewing2020}
\item[Mutational Load]: A measure of the total number of mutations present in the tumour genome.
\item[CNV Load]: Copy-number variation, i.e. deviation the the normal diploid cell compliment.
% \item[Donor survival time]: (should probably remove).
\end{description}

The dendrogram clustering of metasamples (columns, top of figures) gives insights into the similarity of each metasample component, while the clustering of patients (rows, left of figures) offers insights into sub-populations of patients and how (in the case of AOCS only) how these relate to metadata variables such as cellularity and mutational load.   These topics will be taken up in the following discussion section.

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{\plotspath{HeatMaps/clustered_heatmap_TCGA_TCGA.pdf}} 
\caption{Heatmap of the metasamples matrix for the TCGA dataset, computed from metagenes factorized also TCGA.   Clustering dendrograms (based on a correlation distance) are shown between the metasamples (columns) and patients (rows).}
\label{fig-heatmap-TCGA-TCGA}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{\plotspath{HeatMaps/clustered_heatmap_TCGA_AOCS.pdf}} 
\caption{Heatmap of the metasamples matrix for the AOCS dataset, computed from metagenes factorized from the TCGA dataset.   Several patient patient metadata variables are shown in columns to the left.  (ToDo: key for metadata colouring).}
\label{fig-heatmap-TCGA-OACS}
\end{center}
\end{figure}

\clearpage

\section{Discussion}

One valuable lesson from this work is the importance of accounting for sampling error, by bootstrapping,  when considering the stability of factorizations and thence the selection of factorization rank.    Many authors comment on the variability of NMF and ICA factorization due to algorithm initialisation, but neglect sampling error which we have shown to be far more significant.   
These considerations lead to the selection of much lower ranks (fewer metagenes) than would otherwise be the case.

Considering the biological significance of each of the 11 extracted metagenes:
\begin{description}
%
\item[NMF 1-of-3]: this is related to the cellular structures and processes of the extra-cellular matrix (ECM), that is the proteins such as collagens which mediate the three-dimensional organisation of cells in a tissues.  ECM molecular composition will vary substantially between tissue types, but is also known to play a part in many disease processes \cite{Theocharis2019}.  Thus, this metagene may simply reflect heterogeneity of tissues in the biopsied sample, or may have some deeper disease related significance.
%
\item[NMF 2-of-3]: here we see enrichment of genes relating to the ribosomal subunit and the processes of RNA binding, implying perhaps a link with assembly of the ribosomal RNA-protein complex.   Ribosomes are known to have a role in carcinogenesis, by dysregulation the RNA $\rightarrow$ translation or mutations in ribosomal subunits impacting on cellular metabolism \cite{Sulima2017}.
%
\item[NMF 3-of-3]: no significant biological enrichment found.
%
\item[ICA 1-of-5 and ICA 2-of-5]: no significant biological enrichment found.
%
\item[ICA 3-of-5]: this component relates to processes of multicellular / extracellular organisation, and the ECM.  It is therefore similar to NMF 1-of-3 above.
\item[ICA 4-of-5]: this component is seen to relate to the ribosomal subunit (as NMF 2-of-3), additionally membrane proteins in the respiratory complex, mitochondrial and the NADH dehydrogenase complex.  Processes of organonitrogen biosynthesis are also highlighted. 
%
\item[ICA 5-of-5]: processes relating to regulation of immune response are enriched in this component, featuring genes from the major histocompatibility complex group, in particular the HLA- genes, allowing the immune system to recognise self from non-self.  GO terms relating to the endoplasmic reticulum (ER) are also highlighted.  It may be that this component is mainly sensitive to the immunohistochemical signature of patients, and therefore not of clinical interest.
%
\item[PCA 1-of-3]: there is mention of ECM related terms, as for ICA 3-of-5 and NMF 1-of-3.   However, there are also terms relating to the regulation of angiogenesis.  It is known that tumours have a need for increased blood supply, and that expression level of factors proting angigenesis are associated with aggresiveness of tumour growth \cite{Nishida2006}.  Two distinct subtypes of ovarian cancer have been identified, defined on the expression of angiogenesis related genes.  Further, these subtypes have been found to inform clinical outcome \cite{Glass2015}.  Thus, this component might contain useful prognostic value.
%
\item[PCA 2-of-3]: no significant biological enrichment found.
%
\item[PCA 3-of-3]: this component has some similarity with ICA 5-of-5, in that refers to regulation of immune processes (HLA- genes) and ER membrane.   However, chemotaxis (cell movement) is also highlighted.
\end{description}

We have noted above that both NMF 1-of-3 and ICA 3-of-5 may be related to tissue type heterogeneity.  Interestingly, we have also earlier noted that heatmap analysis shows these two metagenes to result in closely correlated metasamples in both the TCGA and OACS datasets.  Further, we also observed that these components correlate negatively with cellularity.  This is as one would expect: low cellularity implies a mixture of tumour and non-tumour cell types, resulting in variation of ECM composition. That these two components highlight similar terms should be no surprise -- they have a high Jaccard similarity of 0.39.

Interestingly, the above summary of the gene enrichment results tentatively suggests component PCA 1-of-3 to be most likely to have clinical prognostic value, based on a cursory look at the literature.  This perhaps supports an earlier observation that, while failing to show consistent significance, component PCA-1-of-3 comes closest to demonstrating correlation with patient survival (section \ref{sec-survival-analysis}).

How should we interpret those four components which demonstrate no significant enrichment with biological meaning?

\begin{itemize}
\item Unsupervised approach
\item Importance of taking account of sampling error.
\item Little correlation with survival time - not surprising.
\item Each method has at least one component with no significantly enriched genes for biological impact - could these be technical dimensions?
\item Lookup most extreme two genes for each component.
\end{itemize}

\section{Conclusions}

%\bibliography{OvarianCancer}
%\bibliographystyle{ieeetr}

\section{Appendices}

\subsection{Software libraries and versions}
\label{sec-software-versions}

\end{document}